\begin{thebibliography}{99}

\bibitem{AssHer06}
\newblock H.E. Assmus, R. Herwig, K.H. Cho and O. Wolkenhauer,
\newblock Dynamics of biological systems: Role of systems biology in medical research,
\newblock {\em Exp Rev Molec Diagn}, Vol. 6, pp. 891--902, 2006.

\bibitem{Arn06}
\newblock C.A. Arnaud,
\newblock Systems biology's clinical future.
\newblock {\em Chem Eng News}, Vol. 84, pp. 17--26, 2006.


\bibitem{Andres93}
\newblock T.~H. Andres, W.~C. Hajas, \newblock{Using iterated fractional factorial design to
\newblock  screen parameters in sensitivity analysis of a probabilistic risk assessment
\newblock  model},  \newblock{\em Proceedings of the joint international conference on
\newblock  mathematical methods and supercomputing in nuclear applications, Karlsruhe,
\newblock  Germany}, Vol.~2, pp. 328--337, 1993.

\bibitem{Bertini} D.J. Bates, J.D. Hauenstein, A.J. Sommese, and C.W. Wampler,
 Bertini\texttrademark: Software for Numerical Algebraic Geometry.
\newblock Available at bertini.nd.edu
           with permanent doi: dx.doi.org/10.7274/R0H41PB5.

\bibitem{BertiniBook} D.J. Bates, J.D. Hauenstein, A.J. Sommese, and C.W. Wampler,
 \newblock {\em Numerically solving polynomial systems with Bertini, SIAM}, 2013.

\bibitem{Barthelmann2000}
V.~Barthelmann, E.~Novak, and K.~Ritter, \newblock High dimensional
polynomial interpolation on sparse grids, \newblock{\em Adv. Comput.
Math.}, Vol.~12,  pp.~273-288, 2000.

\bibitem{Bettonvil90}
\newblock B.~Bettonvil, \newblock Detection of important factors by sequential bifurcation, \newblock{\em Tilburg
\newblock  University Press}, Tilburg, 1990.

\bibitem{BunGen03}
S. D. Bungay, P. A. Gentry, and R. D. Gentry. \newblock A
mathematical model of lipid-mediated thrombin generation.
\newblock {\em Mathematical Medicine and Biology}, Vol. 20, pp.105--129, 2003.

\bibitem{BPM}
K. E. Brummel-Ziedins, R. L. Pouliot, K. G. Mann, \newblock Thrombin
generation: phenotypic quantitation \newblock {\em J Thromb
Haemost}, Vol. 2,  pp. 281--288, 2004.

\bibitem{BWGMR}
K. E. Brummel-Ziedins, M. F. Whelihan, M. Gissel, K. G. Mann, G. E.
Rivard, \newblock Thrombin generation and bleeding in haemophilia A.
\newblock {\em Haemophilia}, Vol. 15, pp. 1118-1125, 2009.

\bibitem{Campolongo05}
\newblock F.~Campolongo, J.~Cariboni, A.~Saltelli, W.~Schoutens, \newblock{Enhancing the Morris
\newblock  method}, \newblock{\em Sensitivity analysis of model output, Kenneth M. Hanson and Francois
\newblock  M. Hemez, eds., Los Alamos National Laboratory}.


\bibitem{CNX}
\newblock S. Christley, Q. Nie and X. Xie. \newblock{Incorporating Existing Network
Information into Gene Network Inference}, \newblock{\em PLoS ONE},
Vol. 4, e6799, 2009.

\bibitem{ChaDen09}
\newblock M. S. Chatterjee, W. S. Denney, H. Jing and S. L. Diamond.
\newblock{Systems biology of coagulation initiation: kinetics of thrombin generation in resting
and activated human blood},
\newblock{\em PLoS Computational Biology}, Vol. 6, Issue 9, e1000950, 2009.

\bibitem{Cools93}
\newblock R.~Cools, \newblock Monomial cubature rules since `stroud': A compilation, \newblock{\em J.
\newblock  Comput. Appl. Math.}, Vol.~48, pp.~309--326, 1993.

\bibitem{Cools99}
\newblock R.~Cools, \newblock Monomial cubature rules since `stroud': A compilation - part
2,
\newblock  {\em J. Comput. Appl. Math.}, Vol.~112, pp.~21--27, 1999.


\bibitem{Cotter79}
\newblock S.~C. Cotter, \newblock A screening design for factorial experiments iwth interactions,
\newblock{\em Biometrika}, Vol. 66 , pp.317--320, 1991.

\bibitem{CF} G. Craciun and M. Feinberg, \newblock Multiple Equilibria in Complex Chemical Reaction Networks: I. The Injectivity Property, {\em SIAM Journal on  Applied Mathematics }, Vol. 65, pp.
1526--1546, 2005.

\bibitem{E1}
\newblock B. Engelmann  \newblock Initiation of coagulation by tissue factor
carriers in blood. {\em Blood Cells Mol. Dis. } Vol. 36, pp.
188--190, 2006.

\bibitem{FogTan05}
A. L. Fogelson and N. Tania.
\newblock Coagulation under flow: the influence of flow-mediated transport on the initiation and inhibition of coagulation.
\newblock {\em Pathophysiology of Haemostasis and Thrombosis}, Vol. 34, pp. 91--108, 2005.


\bibitem{Ganapathysubramanian07}
\newblock B.~Ganapathysubramanian and N.~Zabaras, \newblock Sparse grid collocation schemes for
\newblock  stochastic natural convection problems, \newblock{\em J. Comput. Phys.}, Vol.~225, pp.~652-685, 2007.

\bibitem{Gerstner03}
\newblock T.~Gerstner and M.~Griebel, \newblock Dimension-adaptive tensor-product
quadrature,
 \newblock {\em Computing}, Vol.~71, pp.~65-87, 2003.


\bibitem{Griebel98}
\newblock M.~Griebel, \newblock Adaptive sparse grid multilevel methods for elliptic pdes based
\newblock   on finite differences, \newblock {\em Computing}, Vol.~61,  pp.~151--180, 1998.

\bibitem{HRAMMA}
\newblock A. Hamadeh, M.A.J. Roberts, E. August, P.E. McSharry, P.K. Maini,
J.P. Armitage, \newblock A. Papchristodoulou, Feedback control
architecture \& the bacterial chemotaxis network, \newblock {\em
Plos. Comput. Biol.}, Vol. 7, 2011.



\bibitem{HHHLSZ} W. Hao, J.D. Hauenstein, B.
   Hu, Y. Liu, A.J.
   Sommese, and Y.-T. Zhang \newblock Multiple stable
   steady
   states of
   a reaction-diffusion model on zebrafish
   dorsal-ventral
   patterning\newblock{\em Discrete and Continuous
   Dynamical Systems -
   Series S}, Vol. 4, pp. 1413-1428, 2011.

\bibitem{HHHLSZ2}  W. Hao, J.D. Hauenstein, B.
   Hu, Y. Liu, A.J.
   Sommese, and Y.-T. Zhang,\newblock Bifurcation of
   steady-state solutions for a
tumor model with a necrotic core,\newblock{\em Nonlinear Analysis
Series B: Real World Applications}, Vol. 13, pp. 694--709, 2012.

\bibitem{HHHS}  W. Hao, J.D. Hauenstein, B.
   Hu, and A.J.
   Sommese,\newblock A three-dimensional steady-state
   tumor system,\newblock {\em Applied Mathematics and Computation},
   Vol. 218, pp. 2661--2669, 2011.

\bibitem{HM}
I. D. Hewson, P. Makhmaloaf\newblock Management of third molar
removal with a single dose of recombinant Factor IX (BeneFIX) and
local measures in severe haemophilia B.\newblock {\em Aust. Dent.
J.}, Vol. 55, pp. 322--324, 2010.

\bibitem{HocJon02}
 M. F. Hockin, K. C. Jones, S. J. Everse, and K. G. Mann.
\newblock A Model for the Stoichiometric Regulation of Blood
Coagulation.
\newblock {\em J. Bio. Chem.}, Vol. 277, pp. 18322--18333, 2002.

\bibitem{HWL} K. A. Hogan, H. Weiler, S. T. Lord \newblock Mouse models in coagulation. \newblock {\em Thrombosis and
Haemostasis.}, Vol. 87,  pp. 563--74, 2002.

\bibitem{HacSer06} T. M. Hackeng, K. M. Ser$\acute{e}$, G. Tans and J. Rosing.
\newblock Protein S stimulates inhibition of the tissue factor pathway by tissue factor pathway inhibitor.
\newblock {\em PNAS}, Vol. 103, No. 9,  pp. 3106--3111, 2006.

\bibitem{JesBel93}
J. Jesty, E. Beltrami, and G. Willems.
\newblock Mathematical analysis of a proteolytic positive-feedback loop: dependence
of lag time and enzyme yields on the initial conditions and kinetic
parameters.
\newblock {\em Biochemistry}, Vol. 32, pp. 6266--6274, 1993.

\bibitem{JonMan94a}
K. C. Jones and K. G. Mann.
\newblock A Model for the Tissue Factor Pathway to Thrombin. I. An Empirical Study.
\newblock {\em J. Bio. Chem.}, Vol. 269, pp. 23357--23366, 1994.

\bibitem{JonMan94b}
K. C. Jones and K. G. Mann.
\newblock A Model for the Tissue Factor Pathway to Thrombin. II. A Mathematical Simulation.
\newblock {\em J. Bio. Chem.}, Vol. 269, pp. 23367--23373, 1994.

\bibitem{KhaSem89}
M.A. Khanin  and V.V. Semenov.
\newblock A mathematical model of the kinetics of blood coagulation.
\newblock {\em J. Theor. Biol.}, Vol. 136, pp. 127--134, 1989.

%\bibitem{KMWBHS}
%H.C. Kim, C.W. McMillan, G.C. White, G.E. Bergman, M.W. Horton, P.
%Saidi,\newblock Purified factor IX using monoclonal immunoaffinity
%technique: clinical trials in hemophilia B and comparison to
%prothrombin complex concentrates,\newblock{\em Blood}, Vol. 79 ,
%pp,568--575, 1992.


\bibitem{KF01}
A. L. Kuharsky, A. L. Fogelson,
\newblock Surface-mediated control of blood coagulation: the role of binding site densities and platelet
deposition.
\newblock {\em Biophysical Journal}, Vol. 80, pp. 1050-1074, 2001.

\bibitem{LSK}
M. P. Lambert, B. S. Sachais, M. A. Kowalska, \newblock Chemokines
and thrombogenicity. \newblock {\em Thrombosis and Haemostasis.}
Vol. 97, pp.2-9, 2007.

\bibitem{LGWNC}
A. Lander, K. Gokoffski, F. Wan, Q. Nie, and A. Calof, \newblock
Cell Lineages and the Logic of Proliferative Control. \newblock {\em
PLoS Biology}, Vol. 7, e1000015, 2009.

\bibitem{Li} T.Y. Li \newblock Numerical Solution of Polynomial
   Systems by Homotopy Continuation Methods  in
\newblock{\em   Handbook of Numerical Analysis, Volume XI, Special
   Volume:    Foundations of Computational Mathematics},
   F.~Cucker,   ed.,
   North-Holland, pp. 209--304, 2003.

\bibitem{lin}
\newblock G.~Lin and G.~E.~Karniadakis, \newblock Sensitivity analysis and stochastic simulations of
\newblock non-equilibrium plasma flow,  \newblock{\em Int. J. Numer. Meth. Engng}, Vol. 80, pp.~738--766, 2009.

\bibitem{LinAMTAWR}
\newblock G.~Lin and A.~Tartakovsky, \newblock An efficient, high-order probabilistic collocation method
\newblock on sparse grids for three-dimensional flow and solute
\newblock  transport in randomly heterogeneous porous media, \newblock{\em Advances in Water Resources}, Vol.~32, pp.~712--722, 2009.

\bibitem{LinAMTJSC}
\newblock G.~Lin and A.~Tartakovsky, \newblock Numerical studies of three-dimensional stochastic
\newblock  Darcy's equation and stochastic advection-diffusion-dispersion
equation,
\newblock  {\em Journal of Scientific Computing}, Vol.~43, pp.~92-117, 2010.


\bibitem{LoDen05}
K. Lo, S. Denney, and S. L. Diamond.
\newblock Stochastic Modeling of Blood Coagulation Initiation.
\newblock {\em Pathophysiology of Haemostasis and Thrombosis}, Vol. 34, pp. 80--90, 2005.


\bibitem{LZV}{ D. Luan, M. Zai, J. D. Varner}, \newblock Computationally Derived
Points of Fragility of a Human Cascade Are Consistent with Current
Therapeutic Strategies, \newblock{\em PLoS Computational Biology},
Vol. 3, pp. 142-1359, 2007.


\bibitem{xmazabaras08}
\newblock X.~Ma and N.~Zabaras, \newblock An adaptive hierarchical sparse grid collocation
\newblock  algorithm for the solution of stochastic differential equations, \newblock{\em J. Comput. Phys.}, Vol.~228, pp.~3084-3113, 2009.

%\bibitem{MSWHRL}
%C.W. McMillan, S.S. Shapiro, D. Whitehurst, L.W. Hoyer, A. V. Rao,
%J. Lazerson, \newblock The natural history of factor VIII:C
%inhibitors in patients with hemophilia A: a national cooperative
%study. II. Observations on the initial development of factor VIII:C
%inhibitors,\newblock {\em Blood.} Vol. 71. pp.344--348, 1988.


\bibitem{MCNJY}
T. Moore, C.S. Chou, Q. Nie,  N.L. Jeon and T. M. Yi \newblock{
Robust Spatial Sensing of Mating Pheromone Gradients by Yeast
Cells}, \newblock{\em PLoS ONE}, Vol. 3 :e3865, 2008.

\bibitem{Morris91}
\newblock M.~D. Morris, \newblock Factorial sampling plans for preliminary computational
 \newblock experiments, \newblock{\em Technometrics}, Vol. 33, pp.161--174, 1991.

\bibitem{Nem92}
Y. Nemerson,
\newblock The tissue factor pathway of blood coagulation.
\newblock{\em Semin. Hematol}, Vol. 29, pp. 170-176, 1992.

\bibitem{NOVAK96}
E.~Novak and K.~Ritter, \newblock High dimensional integration of
smooth funcions over cubes, \newblock{\em Numer. Math.}, Vol.~75,
pp.~79-97, 1996.

\bibitem{NOVAK99}
E.~Novak and K.~Ritter, \newblock Simple cubature formulas with high
polynomial exactness, \newblock{\em Constr. Approx.}, Vol.~15,
pp.~499-522, 1999.

\bibitem{PKV13}
Y. Pantazis, M. A. Katsoulakis, D.G. Vlachos.
\newblock Parametric sensitivity analysis for biochemical reaction
networks based on pathwise information theory, \newblock{\em BMC Bioinformatics}, Vol. 14, No. 311, pp.
1-19, 2013.

\bibitem{PWSNCM}
R. Paul, R. Wollman, W. T. Silkworth, I. K. Nardi, D. Cimini, A.
Mogilner. \newblock Computer simulations predict that chromosome
movements and rotations accelerate mitotic spindle assembly without
compromising accuracy, \newblock{\em PNAS}, Vol. 106, pp.
15708-1513, 2009.

\bibitem{Rabitz83}
\newblock H.~Rabitz, M.~Kramer, D.~Dacol, Sensitivity analysis in chemical kinetics,
\newblock {\em Annu. Rev. Phys. Chem.} Vol.~34, pp. 419-461, 1983.


\bibitem{Reagan-2003-combustion}
\newblock M.~T. Reagan, H.~N. Njam, R.~G. Ghanem, O.~M. Knio, \newblock{Uncertainty quantification
\newblock  in reacting-flow simulations through non-intrusive spectral projection},
\newblock{\em Comubstion and Flame} Vol. 132, pp. 545--555, 2003.

\bibitem{Reagan-2005-JCK}
\newblock M.~T. Reagan, H.~N. Njam, R.~G. Ghanem, O.~M. Knio, \newblock{Quantifying uncertainty in
\newblock  chemical systems modeling}, \newblock{\em Int. J. Chem. Kin.}, Vol. 37, pp. 368--382, 2005.



\bibitem{Saltelli00}
\newblock A.~Saltelli, K.~Chan, M.~Scott, \newblock{Sensitivity analysis}, \newblock{\em John Wiley \& Sons Ltd,
\newblock  New York}, 2000.

\bibitem{SMOLYAK63}
\newblock S.~Smolyak, \newblock Quadrature and interpolation formulas for tensor products of
\newblock  certain classes of functions, \newblock{\em Soviet Math. Dokl.}, Vol.~4,  pp.~240--243, 1963.

\bibitem{SJKA1}
\newblock O. ~ Sozinova, Y. Jiang, D. Kaiser, and M. Alber, \newblock A Three-Dimensional Model of Myxobacterial Aggregation by
Contact-mediated Interactions, \newblock{\em Proc. Natl. Acad. Sci.}
Vol. 102, 11308-11312, 2005.

\bibitem{SJKA2}
\newblock O. ~ Sozinova, Y. Jiang, D. Kaiser, and M. Alber, \newblock A
Three-Dimensional Modelof Fruiting Body Formation,
\newblock{\em Proc. Natl. Acad. Sci.}, Vol. 103 pp. 17255-17259, 2006.

\bibitem{SW}{A. J. Sommese, C. W. Wampler}, \newblock {\em The Numerical Solution of
Systems of Polynomials Arising in Engineering and Science},
\newblock {\em World Scientific}, 2005.

\bibitem{SCBM}
\newblock P. Sommi, D. Cheerambathur, I. Brust-Mascher, A. Mogilner,
\newblock Actomyosindependent cortical dynamics contributes to the prophase
force-balance in the early Drosophila embryo, \newblock{\em PLoS
One}, Vol. 6,  e18366, 2011.

\bibitem{tatang}
M.~Tatang and G.~McRae, \newblock Direct treatment of uncertainty in
models of reaction and transport,\newblock{\em MIT Tech. Rep.},
1994.

\bibitem{TG}{ M. Thomson, J. Gunawardena}, \newblock The Rational Parameterisation
Theorem for Multisite Post-translational Modification Systems,
\newblock {\em Journal of Theoretical Biology}, Vol. 261, pp.
626-636, 2009.

\bibitem{TPMA}
\newblock M.J. Tindall, S.L. Porter, P.K. Maini, J.P. Armitage, \newblock Modeling
chemotaxis reveals the role of reversed phosphotransfer \& a
bi-functional kinase-phosphatase, \newblock {\em Plos. Comput.
Biol.}, Vol.6, e1000896, 2010.

\bibitem{PSUADE}
C. Tong
\newblock The PSUADE Uncertainty Quantification Project.
Avialable at
\newblock {\em https://computation.llnl.gov/casc/uncertainty$\sim$quantification/}.

\bibitem{VM}
V. C. Veer, K. G. Mann \newblock Regulation of tissue factor
initiated thrombin generation by the stoichiometric inhibitors
tissue factor pathway inhibitor, antithrombin-III, and heparin
cofactor-II,
\newblock {\em J Biol Chem.} Vol. 7, pp. 4367-4377, 1997.

\bibitem{WJKA}
\newblock Y. Wu, Y. Jiang, D. Kaiser and M. Alber, \newblock Periodic reversal of
direction allows Myxobacteria to swarm, \newblock{\em Proc. Natl.
Acad. Sci.} Vol. 106, pp. 1222-1227, 2009.

\bibitem{DBJSH}
D.~Xiu and J.~S. Hesthaven, \newblock High order collocation methods
for differential equations with random inputs, \newblock {\em SIAM
J. Sci. Comput.}, Vol. 27, pp.1118-1139, 2005.

\bibitem{XLMKLCRA}
Z. ~Xu, J. ~Lioi, J. ~ Mu,  M. M. ~Kamocka, X. ~ Liu, D. Z. ~Chen,
E. D. Rosen and M.~ Alber, A Multiscale Model of Venous Thrombus
Formation with Surface-Mediated Control of Blood Coagulation
Cascade, \newblock {\em  Biophysical Journal},  Vol. 98,
pp.1723--1732 2010.


\bibitem{ZarAta01}
V. I. Zarnitsina, F. I. Ataullakhanov, A. I. Lobanov, and O. L.
Morozova.
\newblock Dynamics of spatially nonuniform patterning in the model of blood coagulation.
\newblock {\em Chaos}, Vol. 11, pp. 57--70, 2001.

\bibitem{ZarPok96}
V. I. Zarnitsina, A. V. Pokhilko, and F. I. Ataullakhanov.
\newblock A mathematical model for the spatio-temporal dynamics of intrinsic pathway of blood coagulation. I. The model description.
\newblock {\em Thrombosis Research}, Vol. 84, pp. 225-236, 1996.


\end{thebibliography}
